Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-17 Master Thesis/synopsis.tex
Clarify 5.4
Finish explaining proof of 4.3

\begin{proof}
\begin{enumerate}[(i)]
\item  We have $(K^\text{nr}_P)^\times/n\cong \mathbb{Z}/n\mathbb{Z}$ since $\operatorname{char}k(P) = 0$. The canonical map $K^\times/n \to (K^\text{nr}_P)^\times/n$ can be identified with the map $K^\times /n\to\mathbb{Z}/n\mathbb{Z}$ induced by $v_P$. Now the following commutative diagram gives us the result:
\[ \begin{tikzcd}
0 \ar[d] & 0 \ar[d]\\
U_{P'}/n \ar[r, dashed]\ar[d] & U_P/n\ar[d]\\
K'^\times /n \ar[r, "\varphi_n"]\ar[d] & K^\times/n\ar[d]\\
(K^\text{nr}_P)'^\times /n \ar[r, "\varphi_n"'] &  (K^\text{nr}_P)^\times /n
\end{tikzcd}\]
Note that $(K^\text{nr}_P)' = (K'^{\text{nr}}_{P'})$ as proven in Lemma \ref{3.2}.
\item Consider the field extension $K_P/K$. Without loss of generality, we can assume $K=K_P$. Then $r_P/n$ is an isomorphism and the composition
\[\begin{tikzcd}
k(P)^\times / n\ar[rr, "(r_P/n)^{-1}"] && U_P/n \ar[rr, hook] && K_P^\times/n
\end{tikzcd}\]
is the inflation $\mathrm{H}^1(k(P), \mu_n) \to \mathrm{H}^1(K_P,\mu_n)$. The same holds for $P'$ due to Lemma \ref{3.2}. Therefore, the assertion follows from the commutativity of the following diagram:
\[ \begin{tikzcd}
k(P')^\times/n \ar[r, "\cong"] \ar[d, "\varphi_P/n"'] & \mathrm{H}^1(k(P'),\mu_n) \ar[r, "\mathrm{inf}"] \ar[d]& \mathrm{H}^1(K'_{P'}, \mu_n) \ar[d, "\varphi_n"] \\
k(P)^\times/n \ar[r, "\cong"'] & \mathrm{H}^1(k(P),\mu_n)  \ar[r, "\mathrm{inf}"'] & \mathrm{H}^1(K_{P}, \mu_n)
\end{tikzcd}  \qedhere\]
\end{enumerate}
\end{proof}

@@

\begin{proof}
Let $U = X\setminus S$ and $P_1\in U$ be a closed point. We set:
\[ C_1 = \ker(r_{P_1}|_{K^S}: K^S \to k(P_1)^\times) \]
Since $r_{P_1}|_{k^\times}:k^\times \to k(P_1)^\times$ is injective, we have $C_1\cap k^\times = 1$ and the following map is injective:
\[ C_1 \to\prod_{P\in S}\mathbb{Z},\ x \mapsto (v_P(x))_{P\in S} \]
Therefore, $C_1$ is a free abelian group of finite rank. If $C_1 = 1$, then we set $T = \{P_1\}$ and we are done. Otherwise, let $x\in C_1,\ x\neq 1$. Then there exists an $a\in k^\times$ with the following properties:
\begin{enumerate}[(i)]
\item There exists a point $P_2\in X_0\setminus(S\cup\{P_1\})$ which lies above the prime of $k(x)$ corresponding to the polynomial $x-a\in k[x]$.
\item $a$ is not a root of unity.
\end{enumerate}
We set $C_2 = \ker(r_{P_2}|_{C_1})\subset C_1$. $C_2$ is free abelian of finite rank as well. By requirement $r_{P_2}(x)=a$ and $a\in k^\times$ is not of finite order, thus $x\not\in C_2$. It follows $\operatorname{rank}(C_2) < \operatorname{rank}(C_1)$. By repeating this process until $C_r=1$, we get points $P_2,\ldots,P_r\in U_0\setminus S$, such that the following map is injective:
\[ C_1\to\prod_{i=2}^{r} k(P_i),\ x\mapsto(r_{P_i}(x))_i \]
Finally for $T=\{P_1,\ldots,P_r\}$, the map $r_{S,T}$ is injective. For the second assertion, note that:
\[ \operatorname{coker}(r_{S,T}) \cong\operatorname{coker}\Big(K^S/k^\times \to \prod_{P\in T}k(P)^\times /k^\times \Big) \]
Since $K^S/k^\times$ is finitely generated, Lemma \ref{5.2} (iii) and Lemma \ref{5.3} (ii) show that the latter group is free-by-finite.
\end{proof}

@@

\section{The Proof}

>>>>

\begin{proof}
\begin{enumerate}[(i)]
\item Consider the maximal unramified extension $L = K_P^\text{nr}$ of $K_P$ and let $w$ be the unique extension of $v_P$ in $L$. Now $k(w) = k(v_P)^\text{s}$. Since $w$ is unramified over $v$, $w$ is discrete as well. We have the following exact sequence:
\[ 1 \longrightarrow \mathcal{O}_w^\times/n \longrightarrow L^\times /n\longrightarrow \mathbb{Z}/n\mathbb{Z}\longrightarrow 0 \]
Since $k(w)$ is separably closed, with Hensel's lemma and $\operatorname{char} k(w)=0$ we can see that $\mathcal{O}_w^\times$ is a divisible group, hence we get an isomorphism $L^\times /n\cong \mathbb{Z}/n\mathbb{Z}$ induced by $w$. Therefore, the canonical map $K^\times/n\to L^\times/n$ can be identified with the map $K^\times/n \to\mathbb{Z}/n\mathbb{Z}$ induced by $v_P$. Now the following commutative diagram with exact columns gives us the result:
\[ \begin{tikzcd}
0 \ar[d] & 0 \ar[d]\\
U_{P'}/n \ar[r, dashed]\ar[d] & U_P/n\ar[d]\\
K'^\times /n \ar[r, "\varphi_n"]\ar[d] & K^\times/n\ar[d]\\
L'^\times /n \ar[r, "\varphi_n"'] &  L^\times /n
\end{tikzcd}\]
Note that $(K^\text{nr}_P)' = K'^{\text{nr}}_{P'}$ as proven in Lemma \ref{3.2}.
\item Consider the field extension $K_P/K$. Without loss of generality, we can assume $K=K_P$. Then $r_P/n$ is an isomorphism and the composition
\[\begin{tikzcd}
k(P)^\times / n\ar[rr, "(r_P/n)^{-1}"] && U_P/n \ar[rr, hook] && K_P^\times/n
\end{tikzcd}\]
is the inflation $\mathrm{H}^1(k(P), \mu_n) \to \mathrm{H}^1(K_P,\mu_n)$. The same holds for $P'$ due to Lemma \ref{3.2}. Therefore, the assertion follows from the commutativity of the following diagram:
\[ \begin{tikzcd}
k(P')^\times/n \ar[r, "\cong"] \ar[d, "\varphi_P/n"'] & \mathrm{H}^1(k(P'),\mu_n) \ar[r, "\mathrm{inf}"] \ar[d]& \mathrm{H}^1(K'_{P'}, \mu_n) \ar[d, "\varphi_n"] \\
k(P)^\times/n \ar[r, "\cong"'] & \mathrm{H}^1(k(P),\mu_n)  \ar[r, "\mathrm{inf}"'] & \mathrm{H}^1(K_{P}, \mu_n)
\end{tikzcd}  \qedhere\]
\end{enumerate}
\end{proof}

@@

\begin{proof}
Let $U = X\setminus S$ and $P_1\in U$ be a closed point. We set:
\[ C_1 = \ker(r_{P_1}|_{K^S}: K^S \to k(P_1)^\times) \]
Since $r_{P_1}|_{k^\times}:k^\times \to k(P_1)^\times$ is injective, we have $C_1\cap k^\times = 1$ and the following map is injective:
\[ C_1 \to\prod_{P\in S}\mathbb{Z},\ x \mapsto (v_P(x))_{P\in S} \]
Therefore, $C_1$ is a free abelian group of finite rank. If $C_1 = 1$, then we set $T = \{P_1\}$ and we are done. Otherwise, let $x\in C_1,\ x\neq 1$. Then there exists an $a\in k^\times$ with the following properties:
\begin{enumerate}[(i)]
\item There exists a point $P_2\in X_0\setminus(S\cup\{P_1\})$ which lies above the prime of $k(x)$ corresponding to the polynomial $x-a\in k[x]$.
\item $a$ is not a root of unity.
\end{enumerate}
We set $C_2 = \ker(r_{P_2}|_{C_1})\subset C_1$. $C_2$ is free abelian of finite rank as well. By requirement $r_{P_2}(x)=a$ and $a\in k^\times$ is not of finite order, thus $x\not\in C_2$. It follows $\operatorname{rank}(C_2) < \operatorname{rank}(C_1)$. By repeating this process until $C_r=1$, we get points $P_2,\ldots,P_r\in U_0$, such that the following map is injective:
\[ C_1\to\prod_{i=2}^{r} k(P_i),\ x\mapsto(r_{P_i}(x))_i \]
Finally for $T=\{P_1,\ldots,P_r\}\subset U_0$, the map $r_{S,T}$ is injective. For the second assertion, note the following isomorphism with the snake lemma:
\[ \operatorname{coker}(r_{S,T}) \cong\operatorname{coker}\Big(K^S/k^\times \to \prod_{P\in T}k(P)^\times /k^\times \Big) \]
Since $K^S/k^\times$ is finitely generated, Lemma \ref{5.2} (iii) and Lemma \ref{5.3} (ii) show that the latter group is free-by-finite.
\end{proof}

@@

\section{Proof of Theorem}
